How to Avoid OOTA Without Really Trying

### 2024-09-13 paper/ootavacuous.tex
ootavacuous: Apply September 13 Alan Stern feedback
Suggested-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

All OOTA workers owe a debt to the foundational work in the infamous
``Causality Test Cases''.\footnote{
	\url{http://www.cs.umd.edu/~pugh/java/memoryModel/unifiedProposal/testcases.html}.}
% %%% Eventually cite P3064.

@@

This section formulates OOTA cycles as cycles in sdep $\cup$ rfe instead
of the traditional sdep $\cup$ rf, then shows that sdep is not a strict
function of the source code.

@@

Semantic dependencies form only one type of link in an OOTA cycle.
The other type extends from a given store to a load that returns the
value stored.
It is tempting to argue that the store must precede
the load in global time and combine this with the intuitive notion that
any real C++ implementation must consume global time when computing a
semantic dependency.
This combination suggests that OOTA cycles cannot occur.
The idea has been formalized by defining an OOTA cycle as a cycle
in sdep $\cup$ rf~\cite{PaulEMcKenney2014OOTA},
% This citation credits the formulation to Ali Sezgin, third author
% of N4323 ("Out-of-Thin-Air Execution is Vacuous").
where sdep is the set of semantic dependencies within
each thread and rf is the set of store-to-load ``reads-from'' links,
whether within a thread (rfi) or between threads (rfe).
% @@@ Consider adding citation defining rf, rfi, and rfe.

>>>>

All OOTA workers owe a debt of gratitude to the foundational work led
by William Pugh leading to the infamous Java-language
``Causality Test Cases''.\footnote{
	\url{http://www.cs.umd.edu/~pugh/java/memoryModel/unifiedProposal/testcases.html}.}
% %%% Eventually cite P3064.

@@

This section analyzes the main components of OOTA cycles, namely
semantic dependencies and reads-from links.

@@

Semantic dependencies form only one type of link in an OOTA cycle.
The other type extends from a given store to a load that returns the
value stored.
It is tempting to argue that the store must precede
the load in global time and combine this with the intuitive notion that
any real C++ implementation must consume global time when computing a
semantic dependency.
This combination suggests that OOTA cycles cannot occur.
The idea has been formalized by defining an OOTA cycle as a cycle
in sdep $\cup$ rf~\cite{PaulEMcKenney2014OOTA},
% This citation credits the formulation to Ali Sezgin, third author
% of N4323 ("Out-of-Thin-Air Execution is Vacuous").
where sdep is the set of semantic dependencies within each thread and
rf is the set of store-to-load ``reads-from'' links, whether internally
within a thread (rfi) or externally between threads (rfe).
% @@@ Consider adding citation defining rf, rfi, and rfe.
